 Total Team Spending and Profitability Results in a Higher League Ranking in the English Premier League

### 2024-04-19 paper/paper.qmd
Committing all cahnge made up to 04-19-2024 @1:35pm.

>>>>

```{r}
#| warning: false
#| echo: false
#| label: fig-actual_pts_vs_predicted_pts
#| fig-cap: A double bar graph comparing the actual and predicted points for teams in the English Premier League. Points predicted by a statistical model are plotted alongside the points actually obtained by the teams to evaluate the models accuracy. The teams are arranged on the x-axis in descending order of their actual points, and two sets of bars represent the actual (in blue) and predicted (in red) points for each team, allowing for a direct side-by-side comparison between the models predictions and real-world outcomes.

# Prepare the predicted rankings with points
predicted_table <- analysis_data |>
  arrange(desc(predicted_pts)) |>
  mutate(predicted_rank = row_number(), `Predicted Points` = predicted_pts) |>
  select(team, predicted_rank, `Predicted Points`)

# Prepare the actual rankings with points
actual_table <- analysis_data |>
  arrange(desc(pts)) |>
  mutate(actual_rank = row_number(), `Actual Points` = pts) |>
  select(team, actual_rank, `Actual Points`)

# Merge both tables based on team
result_table <- merge(actual_table, predicted_table, by = "team")

# Reorder the team factor levels based on actual points
result_table$team <- with(result_table, reorder(team, -`Actual Points`))

# Reshape data to long format for ggplot after reordering
long_data <- result_table |>
  gather(key = "Type", value = "Points", `Actual Points`, `Predicted Points`)

# Create the double bar graph using the reshaped and ordered data
ggplot(long_data, aes(x = team, y = Points, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), width = 0.3) +
  scale_fill_manual(values = c("Actual Points" = "blue", "Predicted Points" = "red"),
                    labels = c("Actual Points", "Predicted Points")) +
  labs(title = "Comparison of Actual and Predicted Points for EPL Teams",
       x = "Team",
       y = "Points",
       fill = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  theme(legend.position = "top")
```
